---
title: "Sampling Properties of the Exponential Distribution -- Project Log"
author: "Brian Davies"
date: "02 September 2014"
output: html_document
---

## Project Aim

Explore the sampling properties of an exponential distribution, as stipulated
in the Coursera Inference Course project rubric.

## Simulation setup

We'll do our basic setup here

```{r set_study_parameters, cache = TRUE}
set.seed( 999*42 )
lambda <- 0.2
sampleSize <- 40
nSamples <- 1000
```

To simulate a sample of 40 exponentials with $\lambda=0.2$, run the following:

```{r single_set_of_40_exponentials, cache = TRUE }
rexp( sampleSize, lambda )
mean( rexp(sampleSize, lambda ) )
sampleSet <- matrix( rexp( nSamples * sampleSize, lambda ), ncol = sampleSize )
sampleMeans <- apply( sampleSet, 1, mean )
hist(sampleMeans)
mean(sampleMeans)
sd(sampleMeans)
```

## Sample statistics of the sample means

So now `sampleSet` contains 1000 simulations of the mean of 40 exponentially-distributed
values, we've calculated their means and looked at the mean and standard deviation of
said sample means, which are `r mean(sampleMeans)` and `r sd(sampleMeans)`
respectively.

For an exponential distribution with $\lambda = 0.2$, the mean $\mu$ and standard
devation are $\sigma$ are both equal to $1 / \lambda$, i.e. `r 1/lambda`. The
expectation of the sample mean is simply the distribution mean $\mu$, and the
standard error of the mean is $\sigma / \sqrt n = 1/ \lambda / \sqrt n$ which
is `r 1/lambda/sqrt(sampleSize)`, i.e pretty close to the values estimated
from the sample.

## Check for normality
