---
title: "Sampling Properties of the Exponential Distribution -- Project Log"
author: "Brian Davies"
date: "19 September 2014"
output: pdf_document
---

## Project Aim

Explore the sampling properties of an exponential distribution, as stipulated
in the Coursera Inference Course project rubric.

## Simulation setup

```{r set_study_parameters, cache = TRUE}
set.seed( 999*42 )
lambda <- 0.2
sampleSize <- 40
nSamples <- 1000
```

To simulate a sample of $n=$ `r sampleSize` exponentials with $\lambda=$ `r lambda`,
run the following:

```{r single_set_of_40_exponentials, cache = TRUE }
rexp( sampleSize, lambda )
mean( rexp(sampleSize, lambda ) )
sampleSet <- matrix( rexp( nSamples * sampleSize, lambda ), ncol = sampleSize )
sampleMeans <- apply( sampleSet, 1, mean )
sampleSDs <- apply( sampleSet, 1, sd )
mean(sampleMeans)
sd(sampleMeans)
```

## Sample statistics of the sample means

So now `sampleSet` contains `r nSamples` simulations of the mean of
`r sampleSize` exponentially-distributed values, we've calculated their means
and looked at the mean and standard deviation of
said sample means, which are `r mean(sampleMeans)` and `r sd(sampleMeans)`
respectively.

For an exponential distribution with $\lambda = $ `r lambda`, the mean $\mu$ and standard
deviation are $\sigma$ are both equal to $1 / \lambda = $ `r 1/lambda`. The
expectation of the sample mean is simply the distribution mean $\mu$, and the
standard error of the mean is $\sigma / \sqrt n = 1/ \lambda / \sqrt n$ which
is `r 1/lambda/sqrt(sampleSize)`, i.e pretty close to the values estimated
from the sample (I'm using the term _pretty close_ in its formal statistical
sense here!)

## Check for normality

Everyone seems to like the Shapiro-Wilk test http://en.wikipedia.org/wiki/Shapiro–Wilk_test

```{r test_for_normality, cache = TRUE}
shapiro.test(sampleMeans)
```

The p-value of `r shapiro.test(sampleMeans)$p.value` __exceeds__ 0.05 so we __fail
to reject the null hypothesis__ that the `sampleMeans` vector is normally distributed.

## Confidence interval coverage

I interpret the question as follows: In what fraction of samples of size $n$
does the _true_ mean $\mu$ lie inside the stated confidence interval as estimated
from the sample? To do this we calculate the upper and lower CI limits for
each sample of `r nSamples` values, and see if the true mean $1 / \lambda$ lies
inside that interval.

```{r coverage_test, cache = TRUE}

ciLo <- sampleMeans - 1.96 * sampleSDs / sqrt(sampleSize)
ciHi <- sampleMeans + 1.96 * sampleSDs / sqrt(sampleSize)
insideCI <- ( ( 1 / lambda ) >= ciLo ) & ( ( 1 / lambda ) <= ciHi )
mean(insideCI)
```

__So our confidence interval has `r mean(insideCI) * 100` percent coverage of
the population mean__
